---
date: "`r format(Sys.Date(), '%d %B, %Y')`"
disclaimer: >
  The scientific results and conclusions, as well as any views or opinions 
  expressed herein, are those of the author(s) and do not necessarily reflect   those of NOAA or the Department of Commerce.
output:
  html_document:
    uswebr::uswds_report
params:
   speno: PV2016_3001
   spot_deployid: PV2016_3001_15U2973
   splash_deployid: PV2016_3001_16A0227
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
library(wcUtils)
library(dplyr)
library(lubridate)
library(rgdal)
library(sp)
library(leaflet)
library(argosfilter)
library(ggplot2)
library(ggthemes)
library(viridis)
```

```{r theme538, include=FALSE}
if (Sys.info()['sysname'] == "Linux") {
  theme_fivethirtyeight <- function(...) {
    ggthemes::theme_fivethirtyeight(
      base_size = 26,
      base_family = "Source Sans Pro"
    )
  }
}
if (Sys.info()['sysname'] == "Darwin") {
  theme_fivethirtyeight <- function(...) {
    ggthemes::theme_fivethirtyeight(
      base_size = 14,
      base_family = "Source Sans Pro"
    )
  }
}
```

```{r get-data, include = FALSE}
wc_response <- wcUtils::wcPOST()
spot_dat <- try(wcUtils::wcGetDeployID(wc_response,deployid = params$spot_deployid),
                silent = TRUE)
splash_dat <- try(wcUtils::wcGetDeployID(wc_response, deployid = params$splash_deployid),
                  silent = TRUE)
if(inherits(spot_dat,"try-error")) { 
  spot_timelines <- NULL
  spot_msg_dt <- NULL
  }
if(inherits(splash_dat,"try-error")) { 
  splash_timelines <- NULL
  splash_msg_dt <- NULL
  }

if (!inherits(spot_dat, "try-error")) {
spot_dat$zipfile <- wcUtils::wcGetZip(id = spot_dat$ids)
spot_dat$dat_files <- unzip(spot_dat$zipfile, exdir = tempdir())
spot_histo_file <- spot_dat$dat_files[grep("*-Histos.csv", 
                                            spot_dat$dat_files)]
spot_histos <- wcUtils::read_histos(spot_histo_file)
spot_timelines <- NULL
if(!inherits(spot_histos,"try-error") && nrow(spot_histos$histos) > 1) {
spot_timelines <- wcUtils::tidyTimelines(spot_histos) %>% 
  mutate(datadatetime = datadatetime + lubridate::dhours(11.5),
         solar_hour = lubridate::hour(datadatetime),
         type = 'SPOT')
}
spot_msg_file <- spot_dat$dat_files[grep("*-All.csv",
                                          spot_dat$dat_files)]
spot_msg_dt <- wcUtils::read_allmsg(spot_msg_file) %>% 
      dplyr::select(msg_date) %>% 
      dplyr::mutate(msg_hour = lubridate::floor_date(msg_date,"hour")) %>%
      dplyr::group_by(msg_hour) %>% 
      dplyr::summarise(msg_count=n())
spot_msg_seq <- data.frame(msg_hour = 
                            seq(min(spot_msg_dt$msg_hour),
                                max(spot_msg_dt$msg_hour + lubridate::hours(1)),
                                by='1 hour'))
spot_msg_dt <- spot_msg_seq %>% dplyr::left_join(spot_msg_dt, by='msg_hour') %>% 
      dplyr::mutate(msg_count = ifelse(is.na(msg_count),0,msg_count),
                    type = 'SPOT') %>% 
      dplyr::arrange(msg_hour)
spot_locs_data <- spot_dat$dat_files[grep(paste0("*",spot_dat$df$ptt,"-Locations.csv"), 
                                 spot_dat$dat_files)]
spot_locs_data <- try(read.csv(spot_locs_data))
}

if (!inherits(splash_dat, "try-error")) {
splash_dat$zipfile <- wcUtils::wcGetZip(id = splash_dat$ids)
splash_dat$dat_files <- unzip(splash_dat$zipfile, exdir = tempdir())

splash_histo_file <- splash_dat$dat_files[grep("*-Histos.csv", 
                                            splash_dat$dat_files)]
splash_histos <- try(wcUtils::read_histos(splash_histo_file), 
                     silent = TRUE)
splash_timelines <- NULL
if(!inherits(splash_histos,"try-error") && nrow(splash_histos$histos) > 1) {
splash_timelines <- wcUtils::tidyTimelines(splash_histos) %>% 
  mutate(datadatetime = datadatetime + lubridate::dhours(11.5),
         solar_hour = lubridate::hour(datadatetime),
         type = 'SPLASH')
}
splash_msg_file <- splash_dat$dat_files[grep("*-All.csv",
                                          splash_dat$dat_files)]

splash_msg_dt <- wcUtils::read_allmsg(splash_msg_file) %>% 
      dplyr::select(msg_date) %>% 
      dplyr::mutate(msg_hour = lubridate::floor_date(msg_date,"hour")) %>%
      dplyr::group_by(msg_hour) %>% 
      dplyr::summarise(msg_count=n())
splash_msg_seq <- data.frame(msg_hour = 
                            seq(min(splash_msg_dt$msg_hour),
                                max(splash_msg_dt$msg_hour + lubridate::hours(1)),
                                by='1 hour'))
splash_msg_dt <- splash_msg_seq %>% dplyr::left_join(splash_msg_dt, by='msg_hour') %>% 
      dplyr::mutate(msg_count = ifelse(is.na(msg_count),0,msg_count),
                    type = 'SPLASH') %>% 
      dplyr::arrange(msg_hour)
locs_fastgps <- try(
  splash_dat$dat_files[grep(paste0("*",splash_dat$df$ptt,"-[0-9]+-Locations.csv"), 
                                   splash_dat$dat_files)]
)
splash_locs_data <- splash_dat$dat_files[grep(paste0("*",splash_dat$df$ptt,"-Locations.csv"), 
                                 splash_dat$dat_files)]
locs_fastgps <- try(wcUtils::read_locs(locs_fastgps))
splash_locs_data <- try(wcUtils::read_locs(splash_locs_data))
    
if (!inherits(locs_fastgps, "try-error")) {
      message(paste('including fastgps data for deployid',splash_dat$df$deployid))
      splash_locs_data <- locs_fastgps
} 

# filter by speed only
mfilter <- argosfilter::vmask(splash_locs_data$latitude,
               splash_locs_data$longitude,
               splash_locs_data$date_time,10)
splash_locs_data <- splash_locs_data[which(mfilter=="not"),]

behav_file <- try(splash_dat$dat_files[grep("*-Behavior.csv", splash_dat$dat_files)])
behavior <- wcUtils::read_behav(behav_file) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(surface_time = 100 * (shallow / (shallow + deep)))
}

if(!is.null(spot_msg_dt) && !is.null(splash_msg_dt)) {
msg_dt <- dplyr::bind_rows(splash_msg_dt,spot_msg_dt)
}
if(!is.null(splash_msg_dt) && is.null(spot_msg_dt)) {
  msg_dt <- splash_msg_dt
}
max_date <- max(msg_dt$msg_hour)
```

## Observed Locations for `r params$speno`

Location data are provided by the SPLASH tag adhered to the seal's hair on either the head or back. For many of the tags deployed during this project, both Argos-based locations and FastLoc/GPS locations are available. While most of the seals also received a flipper-mounted SPOT tag, location data from those tags are not included in this map. A very course speed filter (all steps must be result in a speed less than 10 meters per second) has been applied to remove obviously erroneous locations.

```{r leaflet-map, echo=FALSE}

coordinates(splash_locs_data) <- ~longitude+latitude
proj4string(splash_locs_data) <- CRS("+init=epsg:4326")
splash_locs_data <- sp::spTransform(splash_locs_data,CRS("+proj=longlat +lon_wrap=180"))
splash_locs_track <- as(splash_locs_data,"SpatialLines")

esri_wrld_ocean <- 
  "http://services.arcgisonline.com/arcgis/rest/services/Ocean/World_Ocean_Base/MapServer/tile/{z}/{y}/{x}"
    esri_wrld_ocean_ref <- 
      "http://services.arcgisonline.com/arcgis/rest/services/Ocean/World_Ocean_Reference/MapServer/tile/{z}/{y}/{x}"
    esri_wrld_ocean_attr <- 
      "Tiles &copy; Esri &mdash; Sources: GEBCO, NOAA, CHS, OSU, UNH, CSUMB, National Geographic, DeLorme, NAVTEQ, and Esri"
    
pal <- colorFactor(c("#F19320","#762A83","#EE3333"), domain = c("Argos", "FastGPS","User"))

map <- leaflet(splash_locs_data, width="100%") %>%
        addTiles(urlTemplate = esri_wrld_ocean,
                         attribution = esri_wrld_ocean_attr
                         ) %>%
        addTiles(urlTemplate = esri_wrld_ocean_ref,
                 attribution = esri_wrld_ocean_attr) %>% 
        addPolylines(data = splash_locs_track, weight = 2, color = 'black',
                     group = "Trackline") %>%
        addCircleMarkers(
          radius = 3,
          stroke = FALSE,
          color = ~pal(type),
          fillOpacity = 1
        ) %>% 
        addLegend(pal = pal,
                  values = ~type,
                  title = "Location Type",opacity = 1)
map
```

## Haul-out Behavior

Both the SPLASH and SPOT tags provide haul-out timeline data in the form our hourly percent-dry values. Data are presented as separate components in the plot below. Data are transmitted via the Argos satellite system as messages that cover an entire 24-hour UTC day. Due to the buffer and transmission limits on the tags, there may be days missing from the plot. Hopefully, those data will fill in over time.

```{r haul-out-plot, fig.showtext=TRUE, fig.height=6, warning = FALSE, echo=FALSE}
if (!is.null(splash_timelines) && !is.null(spot_timelines)) {
p <- dplyr::bind_rows(splash_timelines,spot_timelines) %>% 
  ggplot(aes(x = lubridate::date(datadatetime), 
                           y = solar_hour,
                           fill = percent_dry)) +
  geom_tile() +
  scale_y_continuous(breaks = c(6,12,18)) +
  scale_fill_viridis(name = "Percent Dry",
                     guide = guide_colorbar(
                       title.position = "bottom",
                       barwidth = 25, barheight = 1 )) +
  facet_wrap(~type, nrow = 2) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ylab("local solar hour") + xlab("") +
  ggtitle(paste(params$speno,"- Haul-out Behavior by Local Solar Hour")) +
  theme(panel.grid.major.x = element_blank()) 
  print(p)
}
if(is.null(spot_timelines) && !is.null(splash_timelines)) {
p <- splash_timelines %>% 
  ggplot(aes(x = lubridate::date(datadatetime), 
                           y = solar_hour,
                           fill = percent_dry)) +
  geom_tile() +
  scale_y_continuous(breaks = c(6,12,18)) +
  scale_fill_viridis(name = "Percent Dry",
                     guide = guide_colorbar(
                       title.position = "bottom",
                       barwidth = 25, barheight = 1 )) +
  facet_wrap(~type, nrow = 2) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ylab("local solar hour") + xlab("") +
  ggtitle(paste(params$speno,"- Haul-out Behavior by Local Solar Hour")) +
  theme(panel.grid.major.x = element_blank()) 
  print(p)
}
if(is.null(splash_timelines) && !is.null(spot_timelines)) {
p <- spot_timelines %>% 
  ggplot(aes(x = lubridate::date(datadatetime), 
                           y = solar_hour,
                           fill = percent_dry)) +
  geom_tile() +
  scale_y_continuous(breaks = c(6,12,18)) +
  scale_fill_viridis(name = "Percent Dry",
                     guide = guide_colorbar(
                       title.position = "bottom",
                       barwidth = 25, barheight = 1 )) +
  facet_wrap(~type, nrow = 2) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ylab("local solar hour") + xlab("") +
  ggtitle(paste(params$speno,"- Haul-out Behavior by Local Solar Hour")) +
  theme(panel.grid.major.x = element_blank()) 
  print(p)
}
if(is.null(splash_timelines) && is.null(spot_timelines)) {
  p <- NULL
}
```

## Dive Behavior

Dive behavior messages provide details on two categories of activity. An individual 'dive' is defined as a consecutive period of depths below 2 meters that also exceeds 10 meters of depth for more than 10 seconds. For each dive, the start time and end time is recorded as well as the maximum depth. 'Surface' activity is defined as a consectutive period less than 10 meters. For each identified 'surface' period, the behavior is further split into depths from 0-4 meters and from 4-10 meters. Each 'surface' record provides the number of seconds spent in each of the depth categories. In the plot below, the 'surface time' represents the percentage of time in each of the depth categories.

```{r dive-behav-plot, fig.showtext=TRUE, fig.height=6, warning = FALSE, echo=FALSE}
  p <-  ggplot() + 
  geom_rect(data=subset(behavior,what=="Surface"),
            aes(xmin=start,xmax=end,ymin=-4,ymax=0,fill=surface_time),
            linetype=0) +
  geom_rect(data=subset(behavior,what=="Surface"),
            aes(xmin=start,xmax=end,ymin=-10,ymax=-4,fill=100-surface_time),
            linetype=0) +
  scale_fill_distiller(palette = "Greens", direction = 1,
                     name = "Surface Time",
                     guide = guide_colorbar(
                       title.position = "bottom",
                       barwidth = 25, barheight = 1 )) +
  geom_rect(data=subset(behavior,what=="Dive"),
            aes(xmin=start,xmax=end,ymin=-2,ymax=-1*depthmax),linetype=0,
            fill = ggthemes::ptol_pal()(3)[1]) +
  #expand_limits(x=max_date) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  xlab("Date (UTC)") + ylab("Depth (m)") + ggtitle(paste(params$speno,"- Dive Behavior")) + theme(strip.text.y = element_text(size = 7))
print(p)
```

## Message Transmission

Examination of messages received can help troubleshoot issues and evaluate performance.

```{r msg-tx-plot,fig.showtext=TRUE, fig.height=6, warning = FALSE, echo=FALSE}
p <- ggplot(msg_dt,aes(x=msg_hour,y=msg_count)) + 
  geom_line() + 
  facet_wrap(~type, nrow=2) + 
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  xlab("Date (UTC)") + ylab("Messages Transmitted/Hour") + 
  ggtitle(paste(params$speno,"- Message Transmission")) +
  theme(strip.text.y = element_text(size = 7))
print(p)
```
